---
title: "P8105 Homework 5"
author: "Christine Lucille Kuryla (clk2162)"
date: "2023-11-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

## Import and clean data

This problem explores the proportion of homicides in several cities that are solved/result in arrest. Let's start by importing, cleaning, and manipulating the data to serve our purposes.

```{r import_clean_data}
homicide_data <- 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) %>% 
  filter(city_state != "Tulsa, AL") 
```

The resulting dataframe `homicide_data` has information on `r colnames(homicide_data)` and has been cleaned. 

## Proportion of unsolved homicides in Baltimore

```{r prop_baltimore}

# Proportion test to estimite the proportion of unsolved homicides in Baltimore

# filter and select relevant data

baltimore <- homicide_data %>%
  filter(city_state == "Baltimore, MD") %>%
  select(resolution) %>%
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolution == "unsolved")
  )

# perform proportion test

baltimore_prop <- 
    prop.test(
    x = baltimore %>% pull(hom_unsolved),
    n = baltimore %>% pull(hom_total)) 

broom::tidy(baltimore_prop) %>% 
  knitr::kable(digits = 3)

```

The proportion of unsolved homicides in Baltimore is  `r baltimore_prop$estimate`. See above table for details. 

## Proportion of unsolved homicides for all cities in dataset

Note, this code is adapted from the solutions. 

```{r}

# Create a df summarizing the total and unsolved number of homicides in all the cities in the dataset

city_homicide_df <- 
  homicide_data %>% 
  select(city_state, disposition, resolution) %>% 
  group_by(city_state) %>% 
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolution == "unsolved")) 

# Run proportion test on all cities and put in a new df

prop_all_cities_results <- 
  city_homicide_df %>% 
  mutate(
    prop_tests = map2(hom_unsolved, hom_total, \(x, y) prop.test(x = x, n = y)),
    tidy_tests = map(prop_tests, broom::tidy)) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high) 

# plot results and CIs

prop_all_cities_results %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


